---
title: "ANOVA (by hand)"
author: "Heather Siart"
date: "11/27/2021"
Partners: "N/A"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
subtitle: Analysis of Environmental Data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require("here")
rope  = read.csv(here("data", "rope.csv"))
summary(rope)
```
```{r}
levels(rope$rope.type)
```

# ANOVA by hand templet
```{r}
rm(list = ls())

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)
levels(rope$rope.type)

n_obs = 121
n_group = 7

ss_tot = ....
df_tot = ....

agg_sq_resids = ....
ss_within = ....
df_within = ....

ss_among = ....
df_among = ....

ms_within = ....
ms_among  = ....

f_ratio = ....
f_pval = ....
```

# Question 1:
Submit the code you used to build your ANOVA by hand

# Question 2:
Examine the conditional boxplot in the Partitioning Variance: Within-Group section of the walkthrough. Based on the figure, do you think there are equal variances among the groups?

# Question 3:
Conduct a Bartlett test to assess the homogeneity of variances of the percent cut among the rope type groups. Report the p-value.

# Question 4:
Given your graphical assessment (question 2) and the Bartlett test, do you think an ANOVA-type analysis is appropriate on the raw data? Explain why or why not.

# Question 5:
Which rope type is the base case?

Blaze is the base case.

# Question 6:
What is the mean percent cut of the base case rope? Show your calculation using value(s) from the model coefficient table.

# Question 7:
What is the mean percent cut rope type XTC? Show your calculation using value(s) from the model coefficient table.
