---
title: "final_R_reference"
author: "Heather Siart"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}


## Loading Data and Packages
```{r}
require("here")
require(palmerpenguins)
delomys = read.csv(here("data", "delomys.csv"))
ginkgo = read.csv(here("data", "ginkgo_data_2021.csv"))
```

## Data Structures


### C()

The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type.
* I can’t combine character and numeric types in the same call to c()

Here’s two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

I can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec
```

```{r}
## The print() function accomplishes the same task:
print(char_vec)
```

### length()

You can use length() to get or set the length of vectors and factors, and other R objects where a method has been defined.

```{r}
length(ginkgo$max_width)
```

### matrix()

Makes a matrix from a given set of values

```{r}
mat = matrix(1:28, nrow = 4, ncol = 4)
print(mat)
```

### data.frame()
 
Tightly coupled collections of variables which share many of the properties of matrices an of lists, used as the fundamental data structure by most of R's modeling software

```{r}
cat = data.frame(1:10)
print(cat)
```

### nrow()

returns the number of rows in a vector, array, or data frame

nrow(x)

x = a vector, array, data frame, or NULL

```{r}
nrow(mat)
```

### ncol()

returns the number of columns in a vector, array, or data frame

xcol(x)

x = a vector, array, data frame, or NULL

```{r}
ncol(mat)

dat = matrix(1:49, nrow = 7, ncol = 7)
print(dat)
```

### dim()

Dimensions of an object

dim(x)

dim(x) <- value

x = an R object, for example a matrix, array or data frame

value = for a default nethod, either NULL or a numeric vector, which is coerced to integers (by truncation)

```{r}
dim(dat)
```

## Subsetting

### $ 

Subset a data frame by name: select one of the columns in the ginkgo data

```{r}
Gnotch = (ginkgo$notch_depth)
```

### [] 

Use subset by position to:

* select first row of the ginkgo data
* select the element in row 2, column 3
* select the 3rd column of the ginkgo data

```{r}
# First row of the ginkgo data
ginkgo[1, ]

# Element in row 2, column 3
ginkgo[2, 3]

# 3rd column of the ginkgo data
ginkgo[ ,3]
```

### subset() 

Use this function to retrieve all the data for Adelie penguins (in the species column) from the peuguins dataset.

```{r}
All_Adelie = subset(penguins, species=="Adelie")
```

## Numerical Data Exploration

### summary()

summary = used to show results of various model fitting functions or objects

```{r}
summary(num_vec)
```


### mean()

mean = the average of the data

mean(x, ...)

x = an R object

```{r}
mean(ginkgo$max_width)
```

### sd()

sd or standard deviation, 

sd(x, na.rm = FALSE)

x = a numeric vector or an R object but not a factor coercible to numeric by as.double(x)

na.rm = logical. Should missing values be removed? (if FALSE, you don't need to add it into the code)

```{r}
sd(ginkgo$max_width, na.rm = FALSE)

sd(ginkgo$max_width, na.rm = TRUE)
```

## Graphical Data Exploration

### Scatterplot: 
Using the ginkgo data, create a scatterplot of max leaf depth (x) and max leaf width (y).

```{r}
plot(ginkgo$max_depth, ginkgo$max_width,
     main = "Max Leaf Depth vs Max Leaf Width for Ginkgo leaves at UMass", 
     xlab = "Leaf Depth", 
     ylab = "Leaf Width",
     col = "thistle3",
     pch = 20,
     cex = 0.60,        
     xlim = c(15, 160),
     ylim = c(15, 140)
     )
```

### plot() required arguments:

* col = Changes the color of the plotted data
* pch = defines the point symbols in the plot
* cex = Changes the size of the data points on the graph
* main = Title of the graph
* xlab = Label the x axis
* ylab = Label the y axis
* xlim = Sets the range limit of the x axis
* ylim = Sets the range limit of the y axis

### hist() 
* Create a histogram of penguin flipper lengths. Required arguments:

* breaks = are the number of bars on your histogram

```{r}
hist(penguins$flipper_length_mm,
     main = "Flipper Length of Penguins",
     xlab = "Length", col = "palevioletred3",
     xlim = c(160, 250), breaks = 10)
```

### boxplot()
You must include two examples using the ginkgo data:

* 1. a simple boxplot of ginkgo petiole lengths

```{r}
boxplot(ginkgo$petiole_length,
        main = "Petiole Length of Ginkgo Leave at UMass",
        xlab = "",
        ylab = "Length",
        col = "skyblue3")
```

* 2. conditional boxplot of one of the continuous variables conditioned on the seeds_present column.


```{r}
boxplot(ginkgo$seeds_present == FALSE,
        main = "Ginkgo Trees Without Seeds on Campus",
        ylab = "No Seeds",
        col = "azure4")
```

* Create a 4-panel figure of histograms, arranged in a 2 by 2 grid. You may use any data you like, but each histogram must be different and have appropriate titles and axes. - par() required arguments: - mrow =

```{r}
par(mfrow = c(2, 2))
hist(penguins$bill_length_mm,
     main = "Bill Length mm of Penguins",
        xlab = "Length mm",
        ylab = "Frequency",
        col = "darkgoldenrod1")

hist(penguins$bill_depth_mm,
     main = "Bill Depth mm of Penguins",
        xlab = "Depth (mm)",
        ylab = "Frequency",
        col = "darkorange1")

hist(penguins$flipper_length_mm,
     main = "Flipper Length mm of Penguins",
        xlab = "Length (mm)",
        ylab = "Frequency",
        col = "coral2")

hist(penguins$body_mass_g,
     main = "Body Mass (g) of Penguins",
        xlab = "Mass (g)",
        ylab = "Frequency",
        col = "royalblue4")
```

## Distribution Functions

### dnorm()

dnorm or the normal distribution

dnorm(x, mean = 0, sd = 1, log = FALSE)

x = vector quantiles
mean = vector of means
sd = vector of standard deviations
log = logical; if TRUE probablilities p are given as log(p)

### pnorm()

pnorm or the normal distribution

pnorm(q, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)

q = vector quantiles
mean = vector of means
sd = vector of standard deviations
lower.tail = logical; if TRUE (default), probabilities are P[X <= x] otherwise, P[X >x]
log.p = logical; if TRUE probablilities p are given as log(p)

### qnorm()

qnorm or the normal distribution

qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE
)

p = vector of probabilities
mean = vector of means
sd = vector of standard deviations
lower.tail = logical; if TRUE (default), probabilities are P[X <= x] otherwise, P[X >x]
log = logical; if TRUE probablilities p are given as log(p)

### dbinom()

dbinom or the binomial distribution

dbinom(x, size, prob, log = FALSE)

x = vector of quantiles

### pbinom()

pbinom or the binomial distribution

pbinom(q, size, prob, lower.tail = TRUE, log.p = FALSE)

q = vector of quantiles

### qbinom()

qbinom or the binomial distribution

qbinom(p, size, prob, lower.tail = TRUE, log.p = FALSE)

p = vector of probabilities